---
title: "GEO_Code - Foreign Assistance Example"
author: "J. Dayton"
date: "5/13/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
always_allow_html: yes
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(countrycode)
library(dplyr)
# suppressPackageStartupMessages(library(googleVis))
library(plotly)
library(httr)
library(jsonlite)
library(knitr)
library(tinytex)
library(webshot)

source("custFunc.R")
```

## Get the Activity Data

The Foreign Assistance data is from <https://www.foreignassistance.gov/developers>.
This section takes ~30 minutes (+/- 10 minutes).  Process time output to provide feedback that operation remains in execution.

```{r getData, include = FALSE, warning=FALSE}
##Pick what "Activities" we want to get and set the URL
fundingTypeSel <- c("obligated", "spent") #Activity type
agencySel <- c("African%20Development%20Foundation", "Environmental%20Protection%20Agency", "Federal%20Trade%20Commission", "Inter-American%20Foundation", "Millennium%20Challenge%20Corporation", "Overseas%20Private%20Investment%20Corporation", "Peace%20Corps", "U.S.%20Agency%20for%20International%20Development", "Department%20of%20Agriculture", "Department%20of%20Commerce", "Department%20of%20Defense", "Department%20f%20Energy", "Department%20of%20Health%20and%20Human%20Services", "Department%20of%20Homeland%20Security", "Department%20of%20the%20Interior", "Department%20of%20Justice", "Department%20of%20Labor", "Department%20of%20State", "Department%20of%20Transportation", "Department%20of%20the%20Treasury", "Trade%20and%20Development%20Agency", "Export-Import%20Bank", "Department%20of%20the%20Air%20Force", "Department%20of%20the%20Army",  "Department%20of%20the%20Navy", "National%20Science%20Foundation", "Open%20World%20Leadership%20Center", "United%20States%20Institute%20of%20Peace", "Unknown%20-%20Historical%20Greenbook")
yearSel <- as.character(c(2015:2021)) #For years 2015:2021

dfActivity <- getFAActivities(fundingType = fundingTypeSel, agency = agencySel, 
                              year = yearSel)
dfActivity$cCode <- countrycode(dfActivity$benefitingLocation, origin = "country.name", destination = "genc3c")
```

## Get Planned (Activity) Data

The Foreign Assistance data is from <https://www.foreignassistance.gov/developers>.
This section takes ~30 minutes (+/- 10 minutes).  Process time output to provide feedback that operation remains in execution.

```{r planAct, include=FALSE, warning=FALSE }
##Pick what "Planned" we want to get and set the URL


# "https://fagov-apim-prd.azure-api.us/public/api/public/planned?format=json&fundingType=appropriated&filterType=agency&filterValue=Department%20Of%20Defense&year=2013"

##Pick what "Activities" we want to get and set the URL
fundingTypeSel <- c("requested", "appropriated", "actual") #Planned type


dfPlanned <- getFAPlanned(fundingType = fundingTypeSel, agency = agencySel, 
                          year = yearSel) 
dfPlanned$cCode <- countrycode(dfPlanned$benefitingLocation, origin = "country.name", destination = "genc3c")
```

## Create Plot

Plot created for the 2019 Foreign Assistance Planned activites where US dollars were spent in support of the CWMD sector.
``` {r pltlyPlot, echo=FALSE, results = "asis"}
#Set df to plot
tempYear <- 2019
tempFundingType <- "spent" #c("obligated", "spent")
tempSector <- "Combating Weapons of Mass Destruction (WMD)"

dfTemp <- dfActivity[dfActivity$year == tempYear, ]
dfTemp <- dfTemp[dfTemp$fundingType == tempFundingType, ]
dfTemp <- dfTemp[dfTemp$sector == tempSector, ]
dfTemp$amount <- round(dfTemp$amount / 1000000, 2)

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)
# specify map projection/options
g <- list(
  showframe = TRUE,
  showcoastlines = TRUE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(dfTemp) %>%
   add_trace(z = ~amount, color = ~amount, colors = 'Blues',
    text = ~benefitingLocation, locations = ~cCode,
    marker = list(line = l)) %>%
   colorbar(title = '$M (US)', tickprefix = '$') %>%
   layout(title = "TITLE", geo = g)

fig
```


``` {r pltlyPlot, echo=FALSE, results = "asis"}
#Set df to plot
tempYear <- 2019
tempFundingType <- "spent" #c("obligated", "spent")
tempSector <- "Combating Weapons of Mass Destruction (WMD)"

dfTemp <- dfActivity[dfActivity$year == tempYear, ]
dfTemp <- dfTemp[dfTemp$fundingType == tempFundingType, ]
dfTemp <- dfTemp[dfTemp$sector == tempSector, ]
dfTemp$amount <- round(dfTemp$amount / 1000000, 2)

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)
# specify map projection/options
g <- list(
  showframe = TRUE,
  showcoastlines = TRUE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(dfTemp) %>%
   add_trace(z = ~amount, color = ~amount, colors = 'Blues',
    text = ~benefitingLocation, locations = ~cCode,
    marker = list(line = l)) %>%
   colorbar(title = '$M (US)', tickprefix = '$') %>%
   layout(title = "TITLE", geo = g)

fig
```

### Export Method

```{r}
# tmpFile <- tempfile(fileext = ".png")
# export(fig, file = tmpFile)

#In body:
# ![Caption for the picture.](`r tmpFile`)
```



